Exploratory_Analysis_Josh_Tobey

### 2025-12-09 Exploratory_Analysis_Josh_Tobey.Rmd
Preliminary fitting of GAM.

```{r}
time_lag_1 <- 
time_lag_2 <- 
time_lag_3 <-

weekly_log_spec <- logistic_reg()
weekly_log_fit <- weekly_log_spec |>
  fit(Direction ~ Lag1 + 
      Lag2, 
      family = "binomial", 
      data = Weekly)
weekly_log_fit |>
  pluck("fit") |>
  summary()

@@

```{r}
library(mgcv)

lm_spec <- linear_reg()

gam_formula <- Ferric.chloride..mg.L ~ s(Cationic.Polimer..mg.L) + s(mean_Temp) + s(mean_pH) + s(mean_Conductivity) + s(mean_Turbidity) + s(Temp_lag1) + s(Temp_lag2) + s(pH_lag1) + s(pH_lag2) + s(Cond_lag1) + s(Cond_lag2) + s(Turb_lag1) + s(Turb_lag2)

gam_fit <- gam(gam_formula, data = RWAR_dosing_lag_incdec)
gam_rec <- recipe(Ferric.chloride..mg.L ~ ., data = RWAR_dosing_lag_incdec) %>%
  step_ns(Cationic.Polimer..mg.L, deg_free = 10) %>%
  step_ns(mean_Temp, deg_free = 10) %>%
  step_ns(mean_pH, deg_free = 10) %>%
  step_ns(mean_Conductivity, deg_free = 10) %>%
  step_ns(mean_Turbidity, deg_free = 10) %>%
  step_ns(Temp_lag1, deg_free = 10) %>%
  step_ns(Temp_lag2, deg_free = 10) %>%
  step_ns(pH_lag1, deg_free = 10) %>%
  step_ns(pH_lag2, deg_free = 10) %>%
  step_ns(Cond_lag1, deg_free = 10) %>%
  step_ns(Cond_lag2, deg_free = 10) %>%
  step_ns(Turb_lag1, deg_free = 10) %>%
  step_ns(Turb_lag2, deg_free = 10)

gam_wf <- workflow() |>
  add_model(lm_spec) |>
  add_recipe(gam_rec)

gam_fit <- gam_wf |>
  fit(data = RWAR_dosing_lag_incdec)

gam_fit |> tidy()
gam_fit |> glance()

@@

dose_pred <- augment(gam_fit, new_data = dat_grid_dose)
temp_pred <- augment(gam_fit, new_data = dat_grid_temp)
pH_pred <- augment(gam_fit, new_data = dat_grid_pH)
cond_pred <- augment(gam_fit, new_data = dat_grid_cond)
turb_pred <- augment(gam_fit, new_data = dat_grid_turb)
tempLag1_pred <- augment(gam_fit, new_data = dat_grid_temp_lag1)
tempLag2_pred <- augment(gam_fit, new_data = dat_grid_temp_lag2)
pHLag1_pred <- augment(gam_fit, new_data = dat_grid_pH_lag1)
pHLag2_pred <- augment(gam_fit, new_data = dat_grid_pH_lag2)
condLag1_pred <- augment(gam_fit, new_data = dat_grid_cond_lag1)
condLag2_pred <- augment(gam_fit, new_data = dat_grid_cond_lag2)
turbLag1_pred <- augment(gam_fit, new_data = dat_grid_turb_lag1)
turbLag2_pred <- augment(gam_fit, new_data = dat_grid_turb_lag2)

@@

```{r}
for(pred in c("age", "year", "education")) {
  pred_dat <- dose_pred_gam[, c(pred, ".pred", ".pred_lower", ".pred_upper")] |>
    group_by_at(pred) |>
    summarise(.pred = mean(.pred), .pred_lower = mean(.pred_lower), .pred_upper = mean(.pred_upper))
  dose_dat <- RWAR_dosing_lag_incdec[, c(pred, "Ferric.chloride..mg.L")]

  if(is.numeric(dose_dat |> pull(pred))) {
    p <- dose_dat |>
      ggplot() +
      geom_point(aes_string(pred, "wage"), alpha = .3) +
      geom_line(aes_string(pred, ".pred"), data = pred_dat, colour = "darkgreen") +
      geom_line(aes_string(pred, ".pred_lower"), data = pred_dat, colour = "darkgreen", lty = 2) +
      geom_line(aes_string(pred, ".pred_upper"), data = pred_dat, colour = "darkgreen", lty = 2)
  } else {
    p <- dose_dat |>
      ggplot() +
      geom_jitter(aes_string(pred, "wage"), alpha = .3) +
      geom_point(aes_string(pred, ".pred"), data = pred_dat, colour = "darkgreen") +
      geom_errorbar(aes_string(pred, ymin = ".pred_lower", ymax = ".pred_upper"), data = pred_dat, colour = "darkgreen")
  }
  print(p)
}
```

>>>>

```{r}
weekly_log_spec <- logistic_reg()
weekly_log_fit <- weekly_log_spec |>
  fit(Direction ~ Lag1 + 
      Lag2, 
      family = "binomial", 
      data = Weekly)
weekly_log_fit |>
  pluck("fit") |>
  summary()

@@

```{r}
library(mgcv)
library(parsnip)

gam_spec <- gen_additive_mod() |> 
  set_engine("mgcv") |> 
  set_mode("regression")
gam_rec <- recipe(FeCl_dose ~ Polymer_dose + mean_Temp + mean_pH + mean_Conductivity +
                 mean_Turbidity + Temp_lag1 + Temp_lag2 +
                 pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2 +
                 Turb_lag1 + Turb_lag2, data = RWAR_dosing_lag_incdec)

gam_wf <- workflow() |>
  add_model(gam_spec) |>
  add_recipe(gam_rec)

gam_fit <- gam(
  FeCl_dose ~ s(Polymer_dose) + s(mean_Temp) + s(mean_pH) + s(mean_Conductivity) +
               s(mean_Turbidity) + s(Temp_lag1) + s(Temp_lag2) +
               s(pH_lag1) + s(pH_lag2) + s(Cond_lag1) + s(Cond_lag2) +
               s(Turb_lag1) + s(Turb_lag2),
  data = RWAR_dosing_lag_incdec
)

gam_fit |> tidy()
gam_fit |> glance()
```
```{r}

@@

pred1 <- predict(gam_fit, newdata = dat_grid_dose, type = "response", se.fit = TRUE)
dose_pred <- dat_grid_dose |> mutate(
  .fitted = pred1$fit,
  .lower = pred1$fit - 2 * pred1$se.fit,
  .upper = pred1$fit + 2 * pred1$se.fit
  )
pred2 <- predict(gam_fit, newdata = dat_grid_temp, type = "response", se.fit = TRUE)
temp_pred <- dat_grid_temp |> mutate(
  .fitted = pred2$fit,
  .lower = pred2$fit - 2 * pred2$se.fit,
  .upper = pred2$fit + 2 * pred2$se.fit
)
pH_pred <- augment(gam_fit, new_data = dat_grid_pH)
cond_pred <- augment(gam_fit, new_data = dat_grid_cond)
turb_pred <- augment(gam_fit, new_data = dat_grid_turb)
tempLag1_pred <- augment(gam_fit, new_data = dat_grid_temp_lag1)
tempLag2_pred <- augment(gam_fit, new_data = dat_grid_temp_lag2)
pHLag1_pred <- augment(gam_fit, new_data = dat_grid_pH_lag1)
pHLag2_pred <- augment(gam_fit, new_data = dat_grid_pH_lag2)
condLag1_pred <- augment(gam_fit, new_data = dat_grid_cond_lag1)
condLag2_pred <- augment(gam_fit, new_data = dat_grid_cond_lag2)
turbLag1_pred <- augment(gam_fit, new_data = dat_grid_turb_lag1)
turbLag2_pred <- augment(gam_fit, new_data = dat_grid_turb_lag2)

@@

```{r}
ggplot(dose_pred, aes(Polymer_dose, .fitted)) +
  geom_line() + 
  geom_ribbon(aes(ymin = .lower, ymax = .upper), alpha = 0.2)
```
